D.E.S reconstructed lensing potential template and application to delensing and cross correlation

### 2017-01-23 Notes/DES_rec_phi_paper.tex
first

        % Enter your date or \today between curly braces
\begin{abstract}
Goal: test if LSS survey like D.E.S or DESI can improve the reconstruction of the lensing potential that lenses the CMB photons.
This would be crucial to build a template of the B-mode signal coming from the lensing of the primordial E-mode. The lensed B-mode component will be, at multipoles higher than 200, the main contaminants of the primordial B-mode.
Preliminary: CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient with cmb lensing potential constant over $\ell$) of 0.8. Both D.E.S and DESI are significantly worse, $\rho_{\text{eff}}<0.6$. DESI is slightly better than D.E.S. To delens we need the bulk of the redshift distribution to follow the CMB kernel, having a few outliers at redshift $z>1.5$ is not enough. We also have to keep in mind the results of \citet{smith:2012}: even a perfect LSS survey can not help too much the delensing process, with the possible exception of futuristic 21 cm.
\end{abstract}

@@

%=======================================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/spectra.pdf}
\caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$.}
\label{fig:corr}
\end{center}
\end{figure}
%=======================================

%=======================================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/clbb_res.pdf}
\caption{Residual lensing B modes power spectrum using different large scale structure.}
\label{fig:bres}
\end{center}
\end{figure}
%=======================================

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/compare_kernel.pdf}
\caption{Comparison of the different kernels used in this analysis. This allow to understand how well and where in redshift space different LSS surveys trace the CMB lensing potential. 
Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
\label{fig:corr}
\end{center}
\end{figure}
As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is modified by lensing as: 
\be
Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d});~~
U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d})
\ee
where $\mathbf{d}$ is the deflection angle directly related to the lensing potential $\phi$.

As a first approximation, the B mode resulting from the lensing of primordial E mode by a convergence field $\kappa$ is:
\be \label{eq}
B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
\ee
where
\be
W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}),
\ee

@@

\subsection{Galaxies contribution at different redshift}\label{sec:galcontrib}

Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$. It can be shown that the optimal way to combine them to obtain the optimal estimator of $\phi$ or, in other word maximizing the correlation factor $\rho$ is:
\beqn
I 		&=&	\sum_{i}c^{i}I^{i} \nonumber \\
c_{i} 	&=& (C^{-1})_{ij}C^{\kappa I^{j}}
\eeqn
where C is the covariance matrix of the LSS tracers.

\be
\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}
\ee

@@

\bibliographystyle{unsrtnat}
\bibliography{DES_rec_phi_paper}

>>>>

        % Enter your date or \today between curly braces
\begin{abstract}
Goal: test if LSS survey like D.E.S or DESI can improve the reconstruction of the lensing potential that lenses the CMB photons.
This would be crucial to build a template of the B-mode signal coming from the lensing of the primordial E-mode. The lensed B-mode component will be, at multipoles higher than 200, the main contaminants of the primordial B-mode.
Preliminary: CIB performs as an equivalent $\rho_{\text{eff}}$  (correlation coefficient with cmb lensing potential constant over $\ell$) of 0.8. Both D.E.S and DESI are significantly worse, $\rho_{\text{eff}}<0.6$. DESI is slightly better than D.E.S. To delens we need the bulk of the redshift distribution to follow the CMB kernel, having a few outliers at redshift $z>1.5$ is not enough. We also have to keep in mind the results of \cite{smith:2012}: even a perfect LSS survey can not help too much the delensing process, with the possible exception of futuristic 21 cm.
\end{abstract}

@@

%=======================================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/spectra.pdf}
\caption{Correlation factor between galaxies survey and CMB lensing potential as a function of the multipole $\ell$. The ``CIB+D.E.S'' factor has been computed using \refeq{rho-combined}}
\end{center}
\label{fig:corr}

\end{figure}

%=======================================

%=======================================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/clbb_res.pdf}
\caption{Residual lensing B modes power spectrum using different large scale structure. \refeq{combined}}
\end{center}
\label{fig:bres}
\end{figure}

%=======================================

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/compare_kernel.pdf}
\caption{Comparison of the different kernels used in this analysis. This allow to understand how well and where in redshift space different LSS surveys trace the CMB lensing potential. 
Redshift distribution of D.E.S galaxies (I suspect this is the benchmark, anyway taken from Giannantionio et al.). DESI taken from their white paper. }
\end{center}
\label{fig:kernel}
\end{figure}
As for the temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is modified by lensing as: 
\be
Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d});~~
U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\mathbf{d})
\ee
where $\mathbf{d}$ is the deflection angle directly related to the lensing potential $\phi$.

As a first approximation, the B mode resulting from the lensing of primordial E mode by a convergence field $\kappa$ is:
\be 
B^{\mathrm{lens}}(\bl) =  \int \frac{d^2 \bl'}{(2 \pi)^2} W(\bl,\bl') E(\ublu) \kappa(\bl - \bl')
\label{eqn:}
\ee
where
\be
W(\bl,\bl') = \frac{2 \bl' \cdot (\bl-\bl')}{|\bl-\bl'|^2} \sin(2\varphi_{\bl,\bl'}),
\ee

@@

\subsection{Galaxies contribution at different redshifts (CIB+D.E.S)}\label{sec:galcontrib}

Let's now assume that we have n different tracers of the gravitational potentials $I_{i}$ with $i\in \{1,..,n\}$. It can be shown that the optimal way to combine them to estimate $\phi$ or, in other word, maximizing the correlation factor $\rho$ is:
\beqn
I 		&=&	\sum_{i}c^{i}I^{i} \nonumber \\
c_{i} 	&=& (C^{-1})_{ij}C^{\kappa I^{j}}
\label{eqn:combined}
\eeqn 
where C is the covariance matrix of the LSS tracers.
The ``effective'' correlation of these combined tracers with gravitational lensing is: 
\be
\rho^{2} = \sum_{i,j}\frac{C^{\kappa i}~(C^{-1})_{ij}~C^{\kappa j}}{C^{\kappa\kappa}}.
\label{eqn:rho-combined}
\ee

@@

\bibliographystyle{plain}
\bibliography{DES_rec_phi_paper}
